The \dealii Library, Version 9.6

### 2024-09-18 9.6/paper.tex
Fix a latex problem I just introduced :-(
Write a missing section.

    Since very early in the history of the library, the
    \texttt{AffineConstraints} class interface required building such
    constraints in multiple steps: First, one declared a degree of
    freedom as constrained; then one added the dependencies one after
    the other (e.g., by adding pairs $(\frac 12, 13)$ and $(\frac 12,
    14)$ in the hanging node example above); then one added
    inhomogeneities (by setting it to $42$ in the Dirichlet example
    above). This piecemeal approach is cumbersome and prevents the
    library from performing certain error checking steps because a
    constraint is not known to be completely built at any given point. The new
    \textt{add\_constraint()} function now allows defining a
    constraint in one step.
  \item \dealii{} internally needs to keep track of the relative
    orientations of the coordinate systems associated with neighboring
    cells, as well as of the coordinate systems of faces and edges
    within the coordinate system of the cells they are part
    of. Historically, different places within the library that had to
    do that had grown their own, independent ways of representating
    this information, much of which was difficult to understand.
    We have now cleaned up this area. Specifically, instead of using three
    Booleans to define relative orientations for hexedra (\textit{orientation},
    \textit{flip}, and \textit{rotation}), we have internally merged these three values into a single variable of type
     \texttt{char} -- i.e., the smallest integer data type that has at
     least three bits to represent the information. Because the way to represent
     orientations is part of the public interface, the change causes
     some incompatibilities in types and function signatures.
  \todo[inline]{DavidW: Please describe some details here. -- When
    done, please remove this comment.}
\item The \texttt{SolverGMRES} class now offers a third orthogonalization
  method, the classical Gram--Schmidt method with delayed orthogonalization
  \cite{Bielich2022}. Furthermore, the solver now specifies the maximal basis
  size of the Arnoldi basis, rather than the number of auxiliary
  vectors. Some changes have also been made to the GMRES and F-GMRES
  implementations, making them use the same underlying kernels as much as possible.
\end{itemize}
%
The changelog lists more than 180 other features and bugfixes.

@@

\todo[inline]{Wolfgang to write. Mention Lazy and TaskResult.}

>>>>

    Since very early in the history of the library, the
    \texttt{AffineConstraints} class interface required building such
    constraints in multiple steps: First, one declared a degree of
    freedom as constrained; then one added the dependencies one after
    the other (e.g., by adding pairs $(\frac 12, 13)$ and $(\frac 12,
    14)$ in the hanging node example above); then one added
    inhomogeneities (by setting it to $42$ in the Dirichlet example
    above). This piecemeal approach is cumbersome and prevents the
    library from performing certain error checking steps because a
    constraint is not known to be completely built at any given point. The new
    \texttt{add\_constraint()} function now allows defining a
    constraint in one step.
  \item \dealii{} internally needs to keep track of the relative
    orientations of the coordinate systems associated with neighboring
    cells, as well as of the coordinate systems of faces and edges
    within the coordinate system of the cells they are part
    of. Historically, different places within the library that had to
    do that had grown their own, independent ways of representating
    this information, much of which was difficult to understand.
    We have now cleaned up this area. Specifically, instead of using three
    Booleans to define relative orientations for hexedra (\textit{orientation},
    \textit{flip}, and \textit{rotation}), we have internally merged these three values into a single variable of type
     \texttt{char} -- i.e., the smallest integer data type that has at
     least three bits to represent the information. Because the way to represent
     orientations is part of the public interface, the change causes
     some incompatibilities in types and function signatures.
  \todo[inline]{DavidW: Please describe some details here. -- When
    done, please remove this comment.}
\item The \texttt{SolverGMRES} class now offers a third orthogonalization
  method, the classical Gram--Schmidt method with delayed orthogonalization
  \cite{Bielich2022}. Furthermore, the solver now specifies the maximal basis
  size of the Arnoldi basis, rather than the number of auxiliary
  vectors. Some changes have also been made to the GMRES and F-GMRES
  implementations, making them use the same underlying kernels as much as possible.
\end{itemize}
%
The changelog lists more than 180 other features and bugfixes.

@@

Over the years, \dealii{} has accumulated many classes and functions
that support modern programming idioms and make it easier to write
code at higher levels of abstraction.

In the current release, we have added two classes to the list of tools
of this kind:
\begin{itemize}
  \item
  \texttt{Lazy<T>} is a class that supports the lazy computation and
  initialization of variables. Its intended use is for member
  variables of classes that are \textit{sometimes} needed, but perhaps
  not for all uses of an object. For example, all finite element classes
  provide interpolation and restriction matrices to support multigrid
  and other algorithms. One could (i) always compute and store these
  matrices in the constructor of the class; or one could (ii) re-compute these
  matrices every time they are requested. The first of these
  approaches costs memory and compute time even though most places
  where one creates a finite element object will not actually query
  these matrices; the second of these approaches is costly in places
  that \textit{do} query these matrices repeatedly because they are re-computed
  every time. \texttt{Lazy<T>} provides a middle ground: It provides an
  abstraction for an object that is initialized upon first use (that
  is, the first time the value is requested), and then stores the
  computed value for cheap use later on.

  (C++ provides functionality via \texttt{std::async} with
  launch policy \texttt{std::launch::deferred} that can achieve
  similar outcomes. But this functionality is more difficult to use than
  \texttt{Lazy<T>} because, among other reasons, the code generating the
  object has to be specified at the place of construction of the
  object holding the result, rather than at the place of use; and
  because the holder object -- \texttt{std::future} -- can only be
  asked once for its computed value.)

  \item \texttt{TaskResult<T>} is a class that represents the outcome
    of a task possibly evaluated on a separate thread. It can be
    thought of as a ``deferred'' result of a computation in that one
    wants to state ``This job needs to be done, do it when convenient,
    and then put the result of the operation into this
    variable''. Accessing the variable then waits for the operation to
    complete, if it has not already. \texttt{TaskResult<T>} allows
    classes to efficiently compute member variables in the background,
    assuming that they may not be needed right away but only later on.

    (Similar to above, the same effect as \texttt{TaskResult<T>} can
    be achieved using \texttt{std::async}, this time using the
    launch policy \texttt{std::launch::async}. This approach suffers
    from the same issue that one can only query the resulting object
    once. Moreover, \texttt{std::async} does not integrate with the
    thread pool that underlies \dealii{}'s approach to parallel
    processing on modern multi-core machines, whereas
    \texttt{TaskResult<T>} does.)
\end{itemize}
